\chapter{Overview}

\todo{Check urls in bibliography (obtained from (very) old comments!!!}

This is a user manual into the TREMOLO-X Molecular Dynamics simulation package.
Should you be interested in an in-depth introduction into the numerics of molecular dynamics simulation, you might want to read \cite{GriebelEng}
\bigbreak
For up-to-date news and other interesting information about the \mbox{TREMOLO-X} software, please visit:
\begin{center}
 \url{www.tremolo-x.com} 
\end{center}

Tremolo-X is a molecular dynamics simulation package. The program consists of two major components, which allow either sequential or parallel computation of the simulation.

Tremolo-X supplies a choice of several common bonded and non-bonded potentials, which can be combined in order to simulate systems which are dependent on intra- and intermolecular forces.

You may choose from a variety of statistical ensembles and boundary conditions, in order to tailor the simulation to your particular settings and needs.

Furthermore, the dynamic simulations may be augmented or replaced with energy optimization routines.

\section{How to read the manual}

This manual begins with a descriptions of the different capabilities (ensembles, integrators, thermostats,...) of the program, and their interdependencies. We tried to add as much mathematical detail as necessary, but as little as possible, however we always indicate a source, where you can find more details, when you are interested beyond the mere usability of the product.
We continue with an in-depth description of the different potentials you can specify in Tremolo-X and finish with a detailed specification the syntax of options in the parameter file.
